Joker: A Unified Interaction Model For Web Customization

### 2021-08-04 paper.tex
Trim example scenario, move implementation to appendix and add demo video

Section~\ref{sec:examples} describes a concrete scenario, showing how
Joker enables a user to complete a useful customization task. In
Section~\ref{sec:implementation}, we outline the implementation of our
formula language and user interface, as well as the algorithms used by
our PBD interface. We evaluate our approach in
Section~\ref{sec:evaluation} by describing a suite of case studies in
which we used Joker to extract and augment a variety of websites in
order to characterize its capabilities and limitations. Joker relates to
existing work not only in end-user web customization, but also in
end-user web scraping and program synthesis, which we discuss in
Section~\ref{sec:related-work}. Finally, we discuss opportunities for
future work in Section~\ref{sec:conclusion}.

@@

Here is an example scenario, illustrated in Figure~\ref{fig:ebay}, and
demonstrated in the video accompanying this paper. Jen is searching for
a karaoke machine on eBay, a shopping website. She wants to use Joker to
sort products by price within a page of search results, a feature not
supported by eBay.

\emph{Extracting Product Titles \& Prices By Demonstration}.
(Figure~\ref{fig:ebay} Part A): Jen initiates Joker through the browser
context menu. As she hovers over the product title, Joker provides two
kinds of live feedback. First, it highlights the titles of all the
corresponding product listings on the page, to indicate how it has
generalized Jen's intent based on her demonstration of a single example
product. Second, a column of the table is populated with the values that
will be extracted, giving a preview of how the extracted data would
look. To commit to this extraction, Jen clicks on the product title. She
repeats a similar process to extract the price.

When she clicks one of the cells in column \texttt{B}, the formula bar
above the table displays the extraction formula that was generated from
her demonstration:

\texttt{=QuerySelector(rowElement,\ "span.s-item\_\_price")}

This formula produces the values in column \texttt{B}. For each row in
the table, Joker has a created a reference to the corresponding DOM
element, accessible through the special identifier \texttt{rowElement}.
The \texttt{QuerySelector} function runs the specified CSS selector
(\texttt{span.s-item\_\_price}) within the row element to extract the
price of each item. While Jen could directly edit this formula, in this
case the values in the table are correct, so there's no need to edit the
formula.

\emph{Sorting Products By Price}. (Figure~\ref{fig:ebay} Part B): Next,
Jen wants to sort the table by price, but she realizes that the column
of prices is a list of strings containing the \texttt{\$} symbol. She
needs to \emph{augment} this raw data by turning the strings into
numbers. In the next column (\texttt{C}), Jen starts typing a formula,
and uses an autocomplete dropdown to find a relevant function that
extracts numeric values from strings:

\texttt{=ExtractNumber(B)}

Now column C contains numeric values, so Jen can sort the table by
price. Because the table is synchronized with the website, the product
listings become sorted as well.

\emph{Extracting Product URL With Formulas} (Figure~\ref{fig:ebay} Part
C): While completing the previous task, Jen realizes there is another
customization she would find useful: prioritizing products for which she
has not yet visited the details page for the listing. To do this, she
must first return to \emph{extracting} more relevant data from the page.

Each product listing links to a page for the specific product, but
because the URL is not visible in the page, it's not possible to
directly extract it by demonstration. However, Jen can still achieve the
goal by directly writing an extraction formula.

Jen has some basic knowledge of HTML, which she can leverage to write
the formula. She opens the browser developer tools, and observes that
the listing title is represented by a link tag with a heading inside:

\begin{verbatim}
<a href="LISTING PAGE URL">
  <h3>LISTING NAME</h3>
</a>
\end{verbatim}

Since there is already a column \texttt{A} in the table representing the
product's title, she can use this as a starting point to write the
formula:

\texttt{=GetAttribute(GetParent(A),\ "href")}

This formula first calls the \texttt{GetParent} function on column
\texttt{A}, traversing up a level from the
\texttt{\textless{}h3\textgreater{}} elements to the
\texttt{\textless{}a\textgreater{}} elements. Then,
\texttt{GetAttribute} extracts the \texttt{href} value containing the
link URL. After running this formula, the table contains a column
\texttt{D} with the URL for each product.

\emph{Sorting Products By Whether They Have Been Visited}
(Figure~\ref{fig:ebay} Part C): After performing that extraction, Jen
can write a final augmentation formula to indicate whether she has
visited the corresponding product page:

\texttt{=Visited(D)}

The \texttt{Visited} function checks whether a URL is present in the
browser history and returns a boolean value represented by a checkbox.
Jen can then sort by this column to put listings that she has not yet
visited at the top of the page.

Using Joker, Jen was able to not only achieve her initial customization
goal to sort the products by price but also perform a customization she
did not plan to do. This was made possible by Joker's unified
interaction model for web customization which enabled her to interleave
extraction and augmentation. While we have described only a single
illustrative scenario in this example, Joker is flexible enough to
support a wide range of other useful customizations and workflows on
various websites, described in more detail in
Section~\ref{sec:evaluation}.

@@

In this section, we describe Joker's formula language in more detail.
Then, we briefly outline the \emph{wrapper induction}
\citep{kushmerick2000} algorithm that Joker's PBD interface uses to
synthesize the row element and column selectors presented in formulas.

@@

The Wildcard customization tool includes a formula language for
augmentation, including operators for basic arithmetic and string
manipulation, as well as more advanced operators that fetch data from
web APIs. As in other tabular interfaces like SIEUFERD \citep{bakke2016}
and Airtable \citep{2021f}, formulas apply to a whole column at a time
rather than a single cell, and can reference other columns by name.
Joker extends this base language with new constructs which enable it to
apply to \emph{data extraction} instead of just augmentation.

We added DOM elements as a data type in the language, alongside strings,
numbers, and booleans. Because the language runs in a JavaScript
interpreter, we simply use native JavaScript values to represent DOM
elements in the language. DOM elements are displayed visually by showing
their inner text contents. They can also be implicitly typecast to
strings for use in other formulas; for example, a string manipulation
formula like \texttt{Substring} can be called on a DOM element value,
and will operate on its text contents.

@@

When users demonstrate a specific column value to extract, Joker must
synthesize a program that reflects the user's general intent. This is an
instance of the \emph{wrapper induction} problem of synthesizing a web
data extraction query from examples. Prior work on this topic
\citep{kushmerick2000, furche2016} prioritizes accuracy and robustness
to future changes, which makes sense for a fully automated system, but
can lead to very complex queries. In our work, we chose to prioritize
the readability of queries by less sophisticated users, so that users
can more easily author queries and repair them when they break. We
implemented a set of heuristics inspired by Vegemite \citep{lin2009} for
wrapper induction, illustrated in Figure~\ref{fig:overview}.

@@

We evaluate our interaction model and tool by describing the results of
our use of Joker to extract data and perform customizations on popular
websites. For the websites on which Joker can be used, we provide the
sequence of interactions needed to achieve the customizations; for the
websites on which Joker fails, we explain the relevant limitations.

\hypertarget{successful-applications}{%
\subsection{Successful Applications}\label{successful-applications}}

We have used Joker to achieve a variety of customizations across many
websites. Table 1 summarizes examples we have found on popular websites.

@@

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

@@

\begin{table*}[]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Website}              & \textbf{Example Customization Achieved by Joker}                                        \\ \hline
eBay, Amazon            & Filter listings by whether they have a "Sponsored" label.                        \\
Amazon, Target & Sort search results by price and rating.                                                \\
Google Scholar                & Filter publications for those whose title contains a user-provided keyword. \\
Reddit, CNN, ABC  & Sort by the read times of articles. Filter already-visited articles.         \\
Weather.com                   & Filter hourly weather to find sunny times of day.                                        \\
Github                        & Sort a user's code repositories by stars to find popular work.                          \\
Postmates, Uber Eats    & Sort restaurants by delivery time and delivery fee.                                     \\ \hline
\end{tabular}
\vspace{8pt}
\caption{Examples of websites that Joker can be used to customize, including extraction and augmentation}
\end{table*}

@@

Mayer et al propose a user interaction model called \emph{Program
Navigation} \citep{mayer2015} which aims to give users another mechanism
beside examples for guiding the generalization process of PBE tools like
FlashExtract \citep{le2014} and FlashFill \citep{harris}. This is
important because demonstrations are an ambiguous specification for
program synthesis \citep{peleg2018}: the set of synthesized programs for
a demonstration can be very large. Joker shares the general idea of
displaying synthesized programs, but only presents the top-ranked
program.

More broadly, Joker's use of PBD to generate editable code embodies Ravi
Chugh's notion of \emph{prodirect manipulation} \citep{chugh2016a},
implemented in Sketch-N-Sketch \citep{chugh2016}, which aims to bridge
the divide between programmatic and direct manipulation.

@@

@@

>>>>

Section~\ref{sec:examples} describes a concrete scenario, showing how
Joker enables a user to complete a useful customization task. In
Section~\ref{sec:implementation}, we briefly describe the implementation
of our formula language and user interface, as well as the algorithms
used by our PBD interface.

We have performed two evaluations of our approach, presented in
Section~\ref{sec:evaluation}. First, we describe a suite of case studies
in which we used Joker to extract and augment a variety of websites in
order to characterize its capabilities and limitations. Second, we
describe a formative user study with five participants, which showed
that users were generally able to use Joker to perform useful extraction
and augmentation tasks, but which also uncovered limitations,
particularly for less experienced users trying to extract data from more
complex websites.

Joker relates to existing work not only in end-user web customization,
but also in end-user web scraping and program synthesis, which we
discuss in Section~\ref{sec:related-work}. Finally, we discuss
opportunities for future work in Section~\ref{sec:conclusion}.

@@

This section describes an example scenario, illustrated in
Figure~\ref{fig:ebay}, and demonstrated in the video accompanying this
paper. Jen is searching for a karaoke machine on eBay, a shopping
website. She wants to use Joker to sort products by price within a page
of search results, a feature not supported by eBay. Because Joker
supports interleaving extraction and augmentation tasks on the fly, Jen
is also able to extract the URLs of products to determine whether they
have been visited and use that to further sort the products. While we
have described only a single illustrative scenario in this example,
Joker is flexible enough to support a wide range of other useful
customizations and workflows on various websites, described in more
detail in Section~\ref{sec:evaluation}.

@@

In this section, we briefly describe Joker's formula language and the
\emph{wrapper induction} \citep{kushmerick2000} algorithm that Joker's
PBD interface uses to synthesize the row element and column selectors
presented in formulas. Figure~\ref{fig:overview} illustrates the entire
process.

@@

The Wildcard customization tool includes a formula language for
augmentation, including operators for basic arithmetic and string
manipulation, as well as more advanced operators that fetch data from
web APIs. As in other tabular interfaces like SIEUFERD \citep{bakke2016}
and Airtable \citep{2021f}, formulas apply to a whole column at a time
rather than a single cell, and can reference other columns by name.
Joker extends this base language with new constructs which enable it to
apply to \emph{data extraction} instead of just augmentation. More
information can be found in Appendix A.

@@

When users demonstrate a specific column value to extract, Joker must
synthesize a program that reflects the user's general intent. This is an
instance of the \emph{wrapper induction} problem of synthesizing a web
data extraction query from examples. Prior work on this topic
\citep{kushmerick2000, furche2016} prioritizes accuracy and robustness
to future changes, which makes sense for a fully automated system, but
can lead to very complex queries. In our work, we chose to prioritize
the readability of queries by less sophisticated users, so that users
can more easily author queries and repair them when they break. We
implemented a set of heuristics inspired by Vegemite \citep{lin2009} for
wrapper induction, outlined in Appendix B.

@@

We evaluate our interaction model and tool in terms of two research
questions:

\emph{RQ1: What kinds of websites can this model operate effectively
on?} We evaluate this with a suite of case studies that demonstrate its
capabilities and limitations.

\emph{RQ2: How are users of different backgrounds able to use the
system?} We evaluate this with a small formative user study.

\hypertarget{case-studies}{%
\subsection{Case Studies}\label{case-studies}}

Our first evaluation describes the results of our use of Joker to
extract data and perform customizations on popular websites.

\hypertarget{successful-applications}{%
\subsubsection{Successful Applications}\label{successful-applications}}

We have used Joker to achieve a variety of customizations across many
websites.

@@

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

@@

\hypertarget{user-study}{%
\subsection{User Study}\label{user-study}}

Our second evaluation reports a small, formative user study we conducted
to understand how users would interact with Joker.

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We recruited 5 participants with varying backgrounds. 3 participants
were familiar with spreadsheet formulas. 3 participants had extensive
web development experience, 1 had a small amount of prior web
development experience, and 1 had no web development experience. 3
participants had previously extracted data from websites.

\hypertarget{protocol}{%
\subsubsection{Protocol}\label{protocol}}

The participants completed 7 web customization tasks across 2 websites.
All participants attempted all the tasks.

First, we asked participants to customize a website with a relatively
simple HTML structure: the MIT EECS course catalog website. All data
\emph{extraction} on this site can be performed with demonstrations
alone in Joker, although augmentation still requires writing formulas.
The specific tasks were the following: 1a) Extract course titles, 1b)
Extract course prerequisites, 1c) Add a column that indicates whether a
course has a prerequisite \& 1d) Add a column that indicates whether a
course has no prerequisites and is offered in the fall term.

Next, we asked participants to customize a website with a more complex
HTML structure: the search results page for the eBay shopping website.
Due to the website's complexity, demonstrations alone are not sufficient
to extract data; users must also directly edit extraction formulas. The
specific tasks were the following: 2a) Extract title from listings of
Apple iPhones for sale, 2b) Extract the listing price for the phone, \&
2c) Create a column that indicates whether a listing for a phone is
sponsored.

Each session was 60 minutes long and conducted over a recorded video
conference. We started each session with a description of Joker and
provided a brief tutorial of its main features on a sample website not
used in the tasks. There was no time limit for completing the tasks.
Users were encouraged to speak aloud as they worked.

Because some of the tasks build on results of previous tasks, we wanted
to ensure all participants made enough progress to gather useful
feedback. Therefore, whenever a participant got stuck for several
minutes, we recorded why they were stuck and then offered hints on how
to proceed (such as suggestions to read formula documentation or open
the browser dev tools). While all participants were able to complete all
tasks with hints, this obviously does not mean they could have completed
the task unassisted. Our goal was not to simply measure whether users
completed the task, but rather to gain qualitative insight into the
barriers they faced.

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

We have categorized our results into the following four groups:

\emph{Unified interaction model.} Most participants took advantage of
the unified interaction model to interleave extraction and augmentation
tasks, rather than performing all extraction up front. For example, on
task 1, most participants extracted the prerequisites by demonstration,
added one or more columns to the table to perform some string operations
on the prerequisites, and then continued on to extract more information
from the web page by demonstration. Furthermore, we hypothesize that in
a less controlled setting, users would be even more likely to interleave
extraction and augmentation, since the task may be less well defined at
the beginning.

One usability issue with the unified model was that participants
sometimes got confused about how their demonstrations would affect the
contents of the table. For example, multiple participants intended to
add a new column by demonstrating an extraction, but instead
accidentally overwrote the contents of an existing column. This poses a
design challenge because the user's demonstrations occur in the website,
so they cannot directly interact with the table while demonstrating;
this suggests that the interface needs to do a better job indicating
where the results of a demonstration will be inserted.

\emph{Extracting simple data.} On the relatively simple MIT course
catalog website, all participants were able to extract the relevant data
from the page within seconds, simply performing demonstrations with a
few clicks. This suggests that when Joker's generalization algorithm
works well, it can be an effective tool for data extraction, even for
users with limited programming experience. P1 said: \emph{``you could
hover and {[}the data{]} was already selected\ldots that was very
nice''}. P3, upon seeing the tutorial for extraction by demonstration,
said \emph{``that's like black magic.''}

\emph{Extracting complex data.} On the more complex eBay website where
demonstration alone was not sufficient, results were more varied. P1,
who had no prior web development experience, struggled to complete the
task, saying that \emph{``looking at HTML is a bit much''}; this
suggests that more work could be done to make the experience usable for
novices. However, users with more web development experience were able
to use the tool to perform complex extractions, such as directly writing
CSS selectors into the formula bar. P2 and P3 both reported that Joker's
live feedback loop was easier to use and faster than other approaches to
web extraction; P3 noted that \emph{``{[}with any other approach{]}, it
would have been slower to specify and slower to validate that I
specified it correctly.''}

It was challenging for some participants to switch between using the
browser's developer tools and the Joker interface when doing complex
extraction tasks. While we chose not to build HTML inspection into the
Joker UI because the browser already provides a very rich set of tools,
users sometimes were not able to tell how elements in the Joker table
corresponded to elements in the browser's element inspector.

\emph{Writing formulas.} In general, participants were able to learn the
formula language by using an autocomplete dropdown with inline
documentation, which we developed as part of the Joker extension. In
some cases, participants were able to immediately construct correct
formulas on the first try; in other cases it took several attempts and
some hints from the moderator to try a relevant function. While better
documentation and error messages could help improve the learnability of
the formula language, we also did not find it surprising that
participants required some time to learn a completely unfamiliar formula
language.

@@

Mayer et al propose a user interaction model called \emph{Program
Navigation} \citep{mayer2015} which aims to give users another mechanism
beside examples for guiding the generalization process of PBE tools like
FlashExtract \citep{le2014} and FlashFill \citep{harris}. This is
important because demonstrations are an ambiguous specification for
program synthesis \citep{peleg2018}: the set of synthesized programs for
a demonstration can be very large. Joker shares the general idea of
displaying synthesized programs, but only presents the top-ranked
program. More broadly, Joker's use of PBD to generate editable code
embodies Ravi Chugh's notion of \emph{prodirect manipulation}
\citep{chugh2016a}, implemented in Sketch-N-Sketch \citep{chugh2016},
which aims to bridge the divide between programmatic and direct
manipulation.

@@

\newpage

@@

\clearpage
\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}

\hypertarget{appendix-a}{%
\subsection*{A: Extraction Formulas}\label{appendix-a}}

Joker adds DOM elements as a data type to Wildcard's formula language, alongside strings,
numbers, and booleans. Because the language runs in a JavaScript
interpreter, we simply use native JavaScript values to represent DOM
elements in the language. DOM elements are displayed visually by showing
their inner text contents. They can also be implicitly typecast to
strings for use in other formulas; for example, a string manipulation
formula like \texttt{Substring} can be called on a DOM element value,
and will operate on its text contents.

\hypertarget{appendix-b}{%
\subsection*{B: Wrapper Induction Algorithm}\label{appendix-b}}

Our wrapper induction algorithm implements a set of heuristics inspired by Vegemite
as described below:

\hypertarget{determining-row-elements}{%
\subsubsection*{Determining Row
Elements}\label{determining-row-elements}}

The user starts by demonstrating an element \(v\), representing a value
that should be in the table. From that demonstration, we must find a set
of \emph{row elements} that represent the rows of the table. We could
naively assume that \(parent(v)\) is the row containing \(v\), but often
\(v\) is deeply nested inside its containing row; we must determine
which ancestor of \(v\) is likely to be the row.

Intuitively, we solve this problem by assuming that all rows share some
similar internal structure. In particular, we expect most rows to
contain a value for the demonstrated column. (If there were no missing
data, we'd expect \emph{all} rows to contain data for this column.)

Formally: assume a function \(select(el, s)\) which runs a CSS selector
that returns the set of elements matching \(s\) within \(el\). We
generate a set of plausible candidates \(P\), consisting of pairs of a
row element and a CSS selector:

\(P = \{ (r, s) \mid r \in ancestors(v) \land select(r, s) = \{v\} \}\)

For each candidate \((r, s) \in P\), we compute a weight function \(w\),
which is based on the number of siblings of \(r\) that have ``similar
structure'', defined by checking whether running \(s\) within the
sibling also returns a unique element.

\(w(r, s) = |\{ r' \mid r' \in siblings(r) \land |select(r', s) | = 1 \}|\)

We then choose the candidate with the highest weight. In case of ties,
the candidate closer to \(v\) in the tree (i.e., lower in the tree)
wins. Given a winning candidate \((r, s)\), the full set of row elements
is \(\{r\} \cup siblings(r)\).

\hypertarget{synthesizing-css-selectors-for-column-values}{%
\subsubsection*{Synthesizing CSS Selectors For Column
Values}\label{synthesizing-css-selectors-for-column-values}}

Once we have determined the row elements, next we must choose a CSS
selector that will be used to identify the demonstrated value within its
row.

Given a demonstrated value \(v\) within a row element \(r\), we generate
two kinds of plausible selectors:

\begin{itemize}
\tightlist
\item
  selectors using CSS classes, which are manual annotations on DOM
  elements added by the website's programmers, typically for styling
  purposes (e.g.~"item\_\_price")
\item
  selectors using positional indexes within the tree, using the
  \texttt{nth-child} CSS selector (e.g.~\texttt{nth-child(2)},
  representing the second child of an element)
\end{itemize}

The minimum criteria for a plausible selector \(s\) is that it uniquely
identifies the value within the row: \(select(r, s) = \{v\}\). But there
may be many plausible selectors, so we must pick a best one.

We first prioritize selectors using classes, because they tend to be
more robust to changes on the website. A single selector can combine
multiple classes, but we prefer using fewer classes when possible. If no
plausible class-based selector can be generated (for example, if the
relevant elements don't have any classes to query), we fall back to
using a positional index selector. This kind of selector can always be
generated regardless of the contents of the page, but tends to be less
accurate and robust.
